### 2019-01-17 Draft/Parts/Abstract.tex

With the boom in the use of agile process model, imperfect implementations of agile are frequently being seen, out of which the problem of inter-team communication in agile teams is one of the most common issue. An agile team by its definition is a group of people who are self-sufficient to bring their responsibilities to closure, the interaction between different teams is thus considered minimal in most projects implementing agile. This causes problems when the integration of end-products of different teams is carried out. In this paper we have taken a detailed look at this problem and how we can mitigate this.