### 2019-01-21 Draft/Parts/Abstract.tex
Make abstract more clear.

%Problem statement
As the complexity of software development increases, companies relying on agile processes are asked to apply agile at large scale.
A typical approach is to run several agile teams in parallel, however, this can conflict with the basic tenets of agile which focus on interaction and flexibility.
%Proposed Solutions
A key aspect in this transition is maintaining quality communication to ensure progress, which can be addressed with appropriate adaptations in process, technology, and architecture.
%Contribution/Method
In this paper, we review the use of agile for large projects, and appropriate adaptations.