### 2019-01-21 Draft/Parts/Conclusion.tex
Minor fixes to random stuff.

\section{Conclusion}\label{sec:conclusion}
As we have seen, inter-team communication can be an obstacle in agile teams working on the same thing as intermediaries in communication and lack of common personal causes things to go haywire if left unchecked.
To reduce this problem a number of small additions can be done to the process as specified but the most important part of the change is a good architecture and design which is available to all the teams and thus provides a common framework to all the teams involved.