### 2019-01-21 Draft/Parts/Conclusion.tex
Finish pass on introduction.

\section{Conclusion}\label{sec:conclusion}
As we have seen, inter-team communication can be an obstacle in agile teams working together, as intermediaries in communication and lack of personal communication causes a distruption in software process.
To reduce this problem a number of small additions can be done to the process as specified but the most important part of the change is a good architecture and design which is available to all the teams and thus provides a common framework to all the teams involved.