### 2019-01-21 Draft/Parts/Conclusion.tex
Better conclusion.

\section{Conclusion}\label{sec:conclusion}
As we have seen, inter-team communication can be a significant obstacle when agile teams work together, as intermediaries in communication and lack of personal communication causes a disruption in the agile process.
To reduce this problem a number of small adaptions can be made to the process to ensure the flow of information.
However, the most important adaption is a well defined design and architecture which is available across teams and enables teams to produce individual system components which can be easily integrated.