### 2019-02-04 Draft/changes.tex
Added PDF on changes.

\documentclass{article}
\usepackage{graphicx}

\begin{document}

%\title{Introduction to \LaTeX{}}
%\author{Author's Name}

%\maketitle

\section{Assignment 1: Peer Review Changes}
Our paper was reviewed by four reviewers, see the following sections for information on their feedback and changes made. The average evaluation was 2.75 on a scale where 3 was a strong accept.

\subsection{V. A. Madaraju}
\begin{itemize}
    \item Overall evaluation: 3 (strong accept)
    \item Reviewer's confidence: 5 (expert)

This reviewer noted that less importance was given to the design and architecture content, however, they did not explicitly request any changes.
No changes were based based on this feedback alone (a later reviewer noted that the paper has some hanging remarks which we took to include the depth of design and architecture, which we improved).

\end{itemize}

\subsection{C. La Place}
\begin{itemize}
    \item Overall evaluation: 3 (strong accept)
    \item Reviewer's confidence: 4 (high)

This reviewer suggested that we should review our paper for any hanging remarks. They specifically indicated the beginning where we state agile is popular, but go to disadvantages instead of justifying the initial statement. We have rectified this by adding several examples of how following an agile process is beneficial. Together with the comments of V. A. Madaraju, this reviewer's comment about hanging remarks motivated us to add additional commentary on the use and importance of design and architecture.

\end{itemize}

\subsection{D. Bhattacharyya}
\begin{itemize}
    \item Overall evaluation: 2 (accept)
    \item Reviewer's confidence: 4 (high)

This reviewer suggested that sections could have been split up more effectively, or perhaps named more effectively. We reviewed our paper but felt that the existing divisions  and section names suited our work.

\end{itemize}

\subsection{J. Drumm}
\begin{itemize}
    \item Overall evaluation: 3 (strong accept)
    \item Reviewer's confidence: 3 (medium)

This reviewer suggested that we review use of the document organization section in the introduction and decide if it is appropriate. We have determined that the description of the paper's organization is appropriate.

\end{itemize}

\end{document}